Cognitive Principles and Individual Differences \\ in Human Syllogistic Reasoning

### 2019-11-12 paper/paper-v3.tex
added authors names for citations, where necessary

Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
various experimental studies have shown that human reasoning systematically deviates 
from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}).
% Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
% However, the proposed approaches have been strongly critized~(cf.~\cite{mcdermott:87}).
% On the other hand, cognitive scientists have extensively investigated and acquired a great expertise
% in human thinking. Recently, several attempts have been made to bridge the gap among these disciplines. 
% % as they were mostly theoretical and did not apply to real case studies.
% However, only recently an attempt was made 
% this expertise was not consulted, i.e. there was no communication among the relevant fields. 
Cognitive theories that try to explain human reasoning need to be evaluated by their cognitive adequacy~\cite{strube:1992} and their performance should be compared to other theories.
Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}. 
In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~\cite{ragni:stolzenburg:2015} as a starting point:
We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested underlying represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.

% Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable, i.e.\ if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions. 
% Only then, the cognitve adequacy of that theory can be compared to others.
% 
% Formal methods can offer the required tools to implement such cognitive theories. This in turn, can allow a new way of communication among the fields, by avoiding ambiguous definitions, showing their formal correspondences and finally providing appropriate evaluation methods with respect to their cognitive adequacy. 
% A major drawback of 
The results of the meta-study on syllogistic reasoning in~\cite{khemlani:2012} show that the answers given by humans do not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
% Possibly one of the reasons could be that none of the suggested cognitive theories so far has accounted for differences among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities. 
% As recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
% This implies that \textit{the} human reasoner does not exist, instead there 
% seem to exist various (human) reasoning clusters.

@@

According to the authors in~\cite{khemlani:2012}, the majority of participants in experimental studies

@@

Here, we focus on a qualitative assessment of the clustering approach proposed in~\cite{declare:2017},
in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by the authors in~\cite{khemlani:2016}.
In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach. Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach presented in~\cite{declare:2017} on non-aggregated data.
Section~\ref{sect:nonag} provides a more adequate characterization of the individuals' reasoning patterns.

@@

Clustering by principles under WCS was first presented in~\cite{declare:2017};
the authors conjectured that reasoning clusters can be expressed through cognitive principles and heuristic strategies. Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning, 
whereas heuristic strategies describe the methods that are applied when
conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
This approach is an extension of the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
% The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
Figure~\ref{overview:system:syllchal} illustrates the components of the system:
First, groups of reasoners were specified based on a set of underlying cognitive principles and heuristic strategies.
After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
Next, the respective models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
 The final answer distribution was specified according to the number of reasoning clusters 
that entailed the conclusion.
Finally, an illustrative representation of these clusters with respect to their distribution was done by means of multinomial process trees~\cite{Riefer1988}.

@@

Consider the following semantic operator~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
\rangle$ be an interpretation. $\Phi_\CalP (I) =
\langle J^\top,J^\bot\rangle$, where
\[
\begin{array}{lcl}
J^\top & = & \{ A \mid A \leftarrow \Body \in \Def(A,\CalP)
\mbox{ and
} \Body \mbox{ is \True under } \langle I^\top,I^\bot \rangle \} \\
J^\bot & = & \{ A \mid \Def(A,\CalP) \neq \emptyset \mbox{ and }\\
&& \rule{7.5mm}{0mm}\Body \mbox{ is \False under } \langle
I^\top,I^\bot \rangle \mbox{ for all }  A \leftarrow \Body \in \Def(A,\CalP)\}.
\end{array}\Shrink
\]
The weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
always has a least model under  {\L}ukasiewicz logic,
which can be computed by the least fixed point of~$\Svlp$~\cite{hk:2009a}. The correspondence to other well-known logic programming approaches, in particular 
to the well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.

@@

\subsubsection{Conditional Representation (\conditionals) and License for Inference (\licenses)}\label{sect:condli}
Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
However, it seems that humans understand them defeasibly, i.e., they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008} and might allow exceptions to the conditional (\licenses). For (\MA yz) this can be 
explicitly expressed by 
\textit{all $y$ that are not abnormal are $z$}.
% \\
% Accordingly, $z(X)$ is conjoint with an abnormality predicate: 
% \[\forall X (y(X) \wedge \neg \Ab(X) \rightarrow z(X)).\] 
By default, we assume that \textit{nothing is abnormal}. % with respect to $X$}:
% \[
% \Ab(X) \leftarrow \bot.
% \]

@@

Furthermore, humans have a pragmatic understanding of quantifiers in the language~\cite{grice1975}.
For instance, in natural language, we normally do not quantify over things that do not exist.
Consequently, \textit{All $z$} implies \textit{Some $z$ exist}.
This is known as the \textit{Gricean implicature} or the \textit{existential import} (\import),
which also corresponds to the Aristotelian interpretation~\cite{parry:1991} and is assumed by several cognitive theories (cf.~\cite{johnsonlaird:1983,Rips1994}). 
%  Likewise,~\cite{stenning:vanlambalgen:2008} have shown that humans
% require existential import for a conditional to be true.
Together with the introduced cognitive principles in Section~\ref{sect:condli}, the logic program representing (\MA yz), $\CalP_{Ayz}$, is as follows:
\begin{align} %{rll@{\hspace{\hfill}}r}
z(X) & \leftarrow  y(X) \wedge \neg \Ab_{yz}(X). 
\tag{\conditionals\ \&\ \licenses} \\
\Ab_{yz}(X) & \leftarrow \cxt(z'(X)). \tag{\licenses}\\
y(o) & \leftarrow \top. \tag{\import}
\end{align}
Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ will be introduced by the (\footnotesize{\textsf{transformation}}) principle in Section~\ref{sect:dneg}.}
Nothing is abnormal, unless the negation of $z$ is true.
% , then $\Ab_{yz}(X)$ is false, for which then, given the first clause in $\CalP_{Ayx}$, $z(o)$ is true, if $y(o)$ is true. Note that, if $z'(X)$ would be written instead of $\cxt(z'(X))$, in the case where $\cxt(z'(o))$ is unknown, $z(o)$ could not be concluded, even if $y(o)$ would be true. This is due to the fact that negation in WCS is understood as strong negation.\footnote{This is different in other logic programming approaches, i.e.\ well-founded semantics.}
% \end{align}
% \]
We obtain \quad $\CalM_{\CalP_{Ayz}} = \langle \{y(o),z(o)\}, \{\Ab_{yz}(o)\} \rangle$.

@@

\subsubsection{Converse Implication (\converse) and \\ Unknown Generalization (\unknownGen)}
Consider the statement \textit{Some y are z.} ({\MI yz}), which in CL, is represented as $\exists X (y(X) \wedge z(X))$
and which is semantically equivalent to $\exists X(z(X) \wedge y(X))$ (\converse).
As shown by~\cite{khemlani:2012}, humans seem to
distinguish between \textit{some y are z} and \textit{all y are z} (\unknownGen).
Accordingly, if we only observe that an object $o$ belongs to $y$ and $z$ then we do not want to conclude both, \textit{some y are z} and \textit{all y are z}.
In order to distinguish between \textit{some y are z} and \textit{all y are z}, we introduce the following principle:  If we know that \textit{some y are z}, then there must not only be an object~$o_1$ which belongs to $y$ and $z$, but there must be another object $o_2$ which belongs to
$y$ and which does not belong to~$z$.
$\CalP_{Iyz}$, the logic program representing (\MI yz), is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\footnotesize\textsf{converse}}). They can be found in~\cite{declare:2017}.}
\begin{align}
z(X) & \leftarrow y(X) \wedge \neg \Ab_{yz}(X). & \tag{\conditionals\ \&\ \licenses}\\
\Ab_{yz}(o_1) & \leftarrow  \bot. & \tag{\unknownGen\ \& \ \licenses}\\
y(o_1) & \leftarrow  \top. & \tag{\import} \\
y(o_2) & \leftarrow  \top. & \tag{\unknownGen} 
\end{align}
Consider the second clause: Different to the universal moods, \ME\ and \MA, $\Ab_{yz}$ is only false for $o_1$, which in turn implies that 
$\Ab_{yz}$ is unknown for all other objects in the program.
We obtain $\CalM_{\CalP_{Iyz}}= \langle \{y(o_1), y(o_2), z(o_1)\}, \{\Ab_{yz}(o_1)\} \rangle$.
One should observe that  $\Ab_{yz}(o_2)$ is an unknown assumption in
$\CalP_{Iyz}$ and,
hence, $\CalM_{\CalP_{Iyz}}(z(o_2)) = \udf$.

@@

As illustration, let us consider \mbox{\MI\ME 2} from the introduction: 
\begin{align}
 \emph{Some bakers are artists. } & \emph{No chemists are bakers.} \tag{\MI\ME 2} 
\end{align}
The majority of participants in experimental studies
concluded \emph{No chemists are artists} (29\;\%), \emph{No valid conclusion} (27\;\%)
and \emph{Some artists are not chemists} (16\;\%).
In order to account for the individual differences among participants, three clusters on principles and two on heuristic strategies have been suggested
in~\cite{declare:2017}. The two heuristic strategies are the atmosphere bias~\cite{woodworth:sells:1935} and the matching bias~\cite{Wetherick1995}.\footnote{
The interested reader is referred to~\cite{declare:2017} for details on their application in syllogistic reasoning. 
}
The three reasoning clusters have been characterized by the following principles:\footnote{{\footnotesize\textsf{basic}} refers to the set of basic principles introduced in Subsection~\ref{sect:basic}.} (1) \basic, \converse\ for \MI, and \abduction,
    (2) \basic, \converse\ for \MI, and \negFailure, and
    (3) \basic, \converse\ for \MI\ and \ME, and \contraposition\ for~\MA.
This approach achieved a match of 92\;\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches in~\cite{khemlani:2012}.

@@

On aggregated data, the clusters performed fairly well with a match of 92\;\%. 
We are interested whether these clusters also correspond to the reasoning patterns observed among individuals. 
For this purpose, we investigated the non-aggregated dataset provided by~\cite{khemlani:2016} consisting of the responses of 20 participants for two weeks,
i.e.\ the participants carried out the task twice, with a break of one week. 
The results are two-fold:
\begin{description}
 \item[Aggregated evaluation] Most of the majority's answers can be explained by the clusters' predictions. In the first week 
 the majority's conclusion for 63 out of 64 pair of syllogistic premises and in the second week
 all majority's conclusions were predicted.
\item[Non-aggregated evaluation] None of the clusters fitted any individual participants' response pattern, i.e.\ none of the participants belonged to a certain cluster.
\end{description}

@@

\cite{khemlani:2016}\todo{R: cite als Konstituent} proposed an approach to model the individual reasoning patterns of the participants, and 
showed that different participants seem to derive different conclusions.
Their approach is built on top of the mReasoner~\cite{khemlani:2013}, an extension of the mental model theory~\cite{johnsonlaird:1983}.
Reasoning depends on the construction and manipulation of mental models, based on 3 fundamental principles: 
(i) represent possibilities, (ii) principle of iconicity, and (iii) principle of dual processes.
 The procedure of mReasoner consists of two steps:
(1) The construction of a small set of models that represents the terms of the premises, and
(2) the call of a deliberative component to search for counterexamples to conclusions.
They\todo{R: oder doch ``Khemlani''? ;)} specified four parameters for the construction of a mental model and evaluated their
predictions based on 2 datasets, week 1 and week 2, for 20 participants.
The parameters for the construction of the mental models are (i) the 
size of a mental model, (ii) the probability of drawing conclusions only from the most common mental models, or from the complete set,
(iii) the probability to search for counter examples,
and (iv) the probability to weaken the conclusion.
The clustering with mReasoner was based on a particular clustering algorithm to determine the optimal number of clusters for their dataset.
The K-means clustering algorithm was used to divide participants into clusters and then
the mReaoner to simulated the clusters by choosing appropriate parameter settings.
\cite{khemlani:2016}\todo{R: cite als Konstituent} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
The approach is novel in the sense that until now no other cognitive theory tried to rigorously account for individual differences in human syllogistic reasoning. However, their approach seems to have two drawbacks. First, some of their 
parameters were of probabilistic nature, i.e.\ there is no explaining factor behind them. Second, the authors did not distinguish between the participants' 
responses in week 1 and week 2, but instead merged the responses. This might give a distorted picture, as participants could
have learned from week 1 to week 2.

@@

The non-aggregated data provided by~\cite{khemlani:2016} included a dataset of 20 participants,
for which the task was carried out twice, with a break of one week. On average, the basic principles explained the results about 10\;\% better than CL.
 Generally, the participants who were strong in CL/ contraposition, drew fewer conclusions according to abduction/ generalization, and vice versa.
 The changes within participant's answers from one to the other week was high, with 52/64 on average. 
 From week 1 to week 2, an increase with respect to CL (19/20, 6\;\%), basic (18/20, 6\;\%), contraposition (15/20, 5\;\%)
 and a decrease on generalization (15/20, 8\;\%) and no change on abduction (10/20, 0\;\%) answers could be observed.

@@

Similar to~\cite{khemlani:2016}, we identified three different clusters: 
The last two columns show the classification of the participants as deliberative, intuitive or intermediate. The participants in the deliberative cluster here corresponds exactly to~\cite{khemlani:2016}. 
Our classification only differs to~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
A possible explanation for this difference, could be that~\cite{khemlani:2016}
did not take into account the participants' changes between week 1 and week 2.

@@

We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by~\cite{khemlani:2016}.
\cite{khemlani:2016}\todo{R: cite als Konstituent} is the only approach we are aware of that has evaluated predictions on human syllogistic reasoning
with respect to the non-aggregated data.
However in contrast to here, they did not distinguish between the participants' responses in week 1 and week 2.

@@

% Static clusters for each participant might not be appropriate, rather, a sequential application of cognitive principles seems to be more likely
For the future, we need to investigate whether the sequential application of the cognitive principles 
can give us more insights on the participants' reasoning pattern. In particular, our hypothesis is that these principles might correspond to some of the reasoning steps within the insight model~\cite{johnsonlaird:1970} for the selection task~\cite{wason:68}. Similarly, a meta-analysis of the individual response patterns for the selection task has been suggested in~\cite{ragni:2017}.
This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning episodes.

>>>>

Even though the main logical formalism of Classical Logic (CL) has successfully been applied to our scientific reasoning,
various experimental studies have shown that human reasoning systematically deviates 
from conclusions that are valid under CL (cf.~\cite{wason:68,byrne:89}).
% Originally, one of the objectives in the field of AI and logic programming was to formalize human reasoning and commonsense reasoning~\cite{McCarthy59,McCarthy98}. 
% However, the proposed approaches have been strongly critized~(cf.~\cite{mcdermott:87}).
% On the other hand, cognitive scientists have extensively investigated and acquired a great expertise
% in human thinking. Recently, several attempts have been made to bridge the gap among these disciplines. 
% % as they were mostly theoretical and did not apply to real case studies.
% However, only recently an attempt was made 
% this expertise was not consulted, i.e. there was no communication among the relevant fields. 
Cognitive theories that try to explain human reasoning need to be evaluated by their cognitive adequacy~\cite{strube:1992} and their performance should be compared to other theories.
Conceptual cognitive adequacy describes to what extent a system corresponds to human conceptual knowledge, whereas inferential cognitive adequacy indicates whether the reasoning process of a system is structured similarly to the way humans reason~\cite{knauff:rauh:renz:1997}. 
In order to determine the required factors that are necessary for the evaluation of an adequate cognitive theory, we take~the suggestion by Ragni and Stolzburg~\cite{ragni:stolzenburg:2015} as a starting point:
We aim for a theory in which (i) all processes are comprehensible, (ii) the suggested underlying represent general cognitive processes, and (iii) the predictions of our theory are evaluated with respect to a wide range of human reasoning tasks.

% Yet, the cognitive adequacy of a theory can only be measured, if that theory is evaluable, i.e.\ if the theory can uniformally account for episodes of human reasoning systematically, according to its underlying assumptions. 
% Only then, the cognitve adequacy of that theory can be compared to others.
% 
% Formal methods can offer the required tools to implement such cognitive theories. This in turn, can allow a new way of communication among the fields, by avoiding ambiguous definitions, showing their formal correspondences and finally providing appropriate evaluation methods with respect to their cognitive adequacy. 
% A major drawback of 
The results of the meta-study on syllogistic reasoning by Khemlani and Johnson-Laird~\cite{khemlani:2012} show that the answers given by humans do not only systematically differ from the valid conclusions under CL, but from any other proposed cognitive theory (cf.~\cite{BJ89,johnsonlaird:1983,Rips1994,Polk1995,chater:oaksford:1999}).
% Possibly one of the reasons could be that none of the suggested cognitive theories so far has accounted for differences among individuals, i.e.\ they only considered the data summed up with respect to the individual answer possibilities. 
% As recent investigations on psychological experiments show (cf.~\cite{khemlani:2016, ragni:2017}), it seems that different humans represent and reason about conditional information differently.
% This implies that \textit{the} human reasoner does not exist, instead there 
% seem to exist various (human) reasoning clusters.

@@

According to Khemlani and Johnson-Laird~\cite{khemlani:2012}, the majority of participants in experimental studies

@@

Here, we focus on a qualitative assessment of the clustering approach proposed in~\cite{declare:2017},
in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by Johnson-Laird and Khemlani~\cite{khemlani:2016}.
In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach. Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach presented in~\cite{declare:2017} on non-aggregated data.
Section~\ref{sect:nonag} provides a more adequate characterization of the individuals' reasoning patterns.

@@

Clustering by principles under WCS was first presented in~\cite{declare:2017};
the authors conjectured that reasoning clusters can be expressed through cognitive principles and heuristic strategies. Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning, 
whereas heuristic strategies describe the methods that are applied when
conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
This approach is an extension of the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches by Khemlani an Johnson-Laird~\cite{khemlani:2012}.
% The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
Figure~\ref{overview:system:syllchal} illustrates the components of the system:
First, groups of reasoners were specified based on a set of underlying cognitive principles and heuristic strategies.
After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
Next, the respective models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
 The final answer distribution was specified according to the number of reasoning clusters 
that entailed the conclusion.
Finally, an illustrative representation of these clusters with respect to their distribution was done by means of multinomial process trees~\cite{Riefer1988}.

@@

Consider the following semantic operator introduced by Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}: Let $I = \langle I^\top,I^\bot
\rangle$ be an interpretation. $\Phi_\CalP (I) =
\langle J^\top,J^\bot\rangle$, where
\[
\begin{array}{lcl}
J^\top & = & \{ A \mid A \leftarrow \Body \in \Def(A,\CalP)
\mbox{ and
} \Body \mbox{ is \True under } \langle I^\top,I^\bot \rangle \} \\
J^\bot & = & \{ A \mid \Def(A,\CalP) \neq \emptyset \mbox{ and }\\
&& \rule{7.5mm}{0mm}\Body \mbox{ is \False under } \langle
I^\top,I^\bot \rangle \mbox{ for all }  A \leftarrow \Body \in \Def(A,\CalP)\}.
\end{array}\Shrink
\]
The weak completion of non-contextual programs\footnote{Non-contextual programs refer programs that do not contain rules in which $\cxt$ appears in the body of any rule, i.e.\ $p = 0$.} 
always has a least model under  {\L}ukasiewicz logic,
which can be computed by the least fixed point of~$\Svlp$~\cite{hk:2009a}. The correspondence to other well-known logic programming approaches, in particular 
to the well-founded semantics has been shown in~\cite{dietz:hoelldobler:wernhard:2014}.

@@

\subsubsection{Conditional Representation (\conditionals) and License for Inference (\licenses)}\label{sect:condli}
Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
However, it seems that humans understand them defeasibly. According to Stenning and van Lambalgen~\cite{stenning:vanlambalgen:2008}, they assume a \textit{license for the inference} and might allow exceptions to the conditional (\licenses). For (\MA yz) this can be 
explicitly expressed by 
\textit{all $y$ that are not abnormal are $z$}.
% \\
% Accordingly, $z(X)$ is conjoint with an abnormality predicate: 
% \[\forall X (y(X) \wedge \neg \Ab(X) \rightarrow z(X)).\] 
By default, we assume that \textit{nothing is abnormal}. % with respect to $X$}:
% \[
% \Ab(X) \leftarrow \bot.
% \]

@@

Furthermore, according to Grice~\cite{grice1975}, humans have a pragmatic understanding of quantifiers in the language.
For instance, in natural language, we normally do not quantify over things that do not exist.
Consequently, \textit{All $z$} implies \textit{Some $z$ exist}.
This is known as the \textit{Gricean implicature} or the \textit{existential import} (\import),
which also corresponds to the Aristotelian interpretation~\cite{parry:1991} and is assumed by several cognitive theories (cf.~\cite{johnsonlaird:1983,Rips1994}). 
%  Likewise,~\cite{stenning:vanlambalgen:2008} have shown that humans
% require existential import for a conditional to be true.
Together with the introduced cognitive principles in Section~\ref{sect:condli}, the logic program representing (\MA yz), $\CalP_{Ayz}$, is as follows:
\begin{align} %{rll@{\hspace{\hfill}}r}
z(X) & \leftarrow  y(X) \wedge \neg \Ab_{yz}(X). 
\tag{\conditionals\ \&\ \licenses} \\
\Ab_{yz}(X) & \leftarrow \cxt(z'(X)). \tag{\licenses}\\
y(o) & \leftarrow \top. \tag{\import}
\end{align}
Note that $z'(X)$ appears within the $\cxt$ operator in the body of the second clause, where $z'$ represents the negation of $z$:\footnote{$z'(X)$ will be introduced by the (\footnotesize{\textsf{transformation}}) principle in Section~\ref{sect:dneg}.}
Nothing is abnormal, unless the negation of $z$ is true.
% , then $\Ab_{yz}(X)$ is false, for which then, given the first clause in $\CalP_{Ayx}$, $z(o)$ is true, if $y(o)$ is true. Note that, if $z'(X)$ would be written instead of $\cxt(z'(X))$, in the case where $\cxt(z'(o))$ is unknown, $z(o)$ could not be concluded, even if $y(o)$ would be true. This is due to the fact that negation in WCS is understood as strong negation.\footnote{This is different in other logic programming approaches, i.e.\ well-founded semantics.}
% \end{align}
% \]
We obtain \quad $\CalM_{\CalP_{Ayz}} = \langle \{y(o),z(o)\}, \{\Ab_{yz}(o)\} \rangle$.

@@

\subsubsection{Converse Implication (\converse) and \\ Unknown Generalization (\unknownGen)}
Consider the statement \textit{Some y are z.} ({\MI yz}), which in CL, is represented as $\exists X (y(X) \wedge z(X))$
and which is semantically equivalent to $\exists X(z(X) \wedge y(X))$ (\converse).
As shown by Khemlani and Johnson-Laird~\cite{khemlani:2012}, humans seem to
distinguish between \textit{some y are z} and \textit{all y are z} (\unknownGen).
Accordingly, if we only observe that an object $o$ belongs to $y$ and $z$ then we do not want to conclude both, \textit{some y are z} and \textit{all y are z}.
In order to distinguish between \textit{some y are z} and \textit{all y are z}, we introduce the following principle:  If we know that \textit{some y are z}, then there must not only be an object~$o_1$ which belongs to $y$ and $z$, but there must be another object $o_2$ which belongs to
$y$ and which does not belong to~$z$.
$\CalP_{Iyz}$, the logic program representing (\MI yz), is as follows:\footnote{For simplicity, here we omit the clauses introduced by ({\footnotesize\textsf{converse}}). They can be found in~\cite{declare:2017}.}
\begin{align}
z(X) & \leftarrow y(X) \wedge \neg \Ab_{yz}(X). & \tag{\conditionals\ \&\ \licenses}\\
\Ab_{yz}(o_1) & \leftarrow  \bot. & \tag{\unknownGen\ \& \ \licenses}\\
y(o_1) & \leftarrow  \top. & \tag{\import} \\
y(o_2) & \leftarrow  \top. & \tag{\unknownGen} 
\end{align}
Consider the second clause: Different to the universal moods, \ME\ and \MA, $\Ab_{yz}$ is only false for $o_1$, which in turn implies that 
$\Ab_{yz}$ is unknown for all other objects in the program.
We obtain $\CalM_{\CalP_{Iyz}}= \langle \{y(o_1), y(o_2), z(o_1)\}, \{\Ab_{yz}(o_1)\} \rangle$.
One should observe that  $\Ab_{yz}(o_2)$ is an unknown assumption in
$\CalP_{Iyz}$ and,
hence, $\CalM_{\CalP_{Iyz}}(z(o_2)) = \udf$.

@@

As illustration, let us consider \mbox{\MI\ME 2} from the introduction: 
\begin{align}
 \emph{Some bakers are artists. } & \emph{No chemists are bakers.} \tag{\MI\ME 2} 
\end{align}
The majority of participants in experimental studies
concluded \emph{No chemists are artists} (29\;\%), \emph{No valid conclusion} (27\;\%)
and \emph{Some artists are not chemists} (16\;\%).
In order to account for the individual differences among participants, three clusters on principles and two on heuristic strategies have been suggested
in~\cite{declare:2017}. The two heuristic strategies are the atmosphere bias~\cite{woodworth:sells:1935} and the matching bias~\cite{Wetherick1995}.\footnote{
The interested reader is referred to~\cite{declare:2017} for details on their application in syllogistic reasoning. 
}
The three reasoning clusters have been characterized by the following principles:\footnote{{\footnotesize\textsf{basic}} refers to the set of basic principles introduced in Subsection~\ref{sect:basic}.} (1) \basic, \converse\ for \MI, and \abduction,
    (2) \basic, \converse\ for \MI, and \negFailure, and
    (3) \basic, \converse\ for \MI\ and \ME, and \contraposition\ for~\MA.
This approach achieved a match of 92\;\% with the aggregated data reported in~\cite{khemlani:2012}, and was the highest result compared to the presented approaches.

@@

On aggregated data, the clusters performed fairly well with a match of 92\;\%. 
We are interested whether these clusters also correspond to the reasoning patterns observed among individuals. 
For this purpose, we investigated the non-aggregated dataset provided by Johnson-Laird and Khemlani~\cite{khemlani:2016} consisting of the responses of 20 participants for two weeks,
i.e.\ the participants carried out the task twice, with a break of one week. 
The results are two-fold:
\begin{description}
 \item[Aggregated evaluation] Most of the majority's answers can be explained by the clusters' predictions. In the first week 
 the majority's conclusion for 63 out of 64 pair of syllogistic premises and in the second week
 all majority's conclusions were predicted.
\item[Non-aggregated evaluation] None of the clusters fitted any individual participants' response pattern, i.e.\ none of the participants belonged to a certain cluster.
\end{description}

@@

Johnson-Laird and Khemlani~\cite{khemlani:2016} proposed an approach to model the individual reasoning patterns of the participants, and 
showed that different participants seem to derive different conclusions.
Their approach is built on top of the mReasoner~\cite{khemlani:2013}, an extension of the mental model theory, first introduced by Johnson-Laird~\cite{johnsonlaird:1983}.
Reasoning depends on the construction and manipulation of mental models, based on 3 fundamental principles: 
(i) represent possibilities, (ii) principle of iconicity, and (iii) principle of dual processes.
 The procedure of mReasoner consists of two steps:
(1) The construction of a small set of models that represents the terms of the premises, and
(2) the call of a deliberative component to search for counterexamples to conclusions.
Johnson-Laird and Khemlani specified four parameters for the construction of a mental model and evaluated their
predictions based on 2 datasets, week 1 and week 2, for 20 participants.
The parameters for the construction of the mental models are (i) the 
size of a mental model, (ii) the probability of drawing conclusions only from the most common mental models, or from the complete set,
(iii) the probability to search for counter examples,
and (iv) the probability to weaken the conclusion.
The clustering with mReasoner was based on a particular clustering algorithm to determine the optimal number of clusters for their dataset.
The K-means clustering algorithm was used to divide participants into clusters and then
the mReaoner to simulated the clusters by choosing appropriate parameter settings.
Johnson-Laird and Khemlani~\cite{khemlani:2016} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
The approach is novel in the sense that until now no other cognitive theory tried to rigorously account for individual differences in human syllogistic reasoning. However, their approach seems to have two drawbacks. First, some of their 
parameters were of probabilistic nature, i.e.\ there is no explaining factor behind them. Second, the authors did not distinguish between the participants' 
responses in week 1 and week 2, but instead merged the responses. This might give a distorted picture, as participants could
have learned from week 1 to week 2.

@@

The non-aggregated data included a dataset of 20 participants,
for which the task was carried out twice, with a break of one week~\cite{khemlani:2016}. On average, the basic principles explained the results about 10\;\% better than CL.
 Generally, the participants who were strong in CL/ contraposition, drew fewer conclusions according to abduction/ generalization, and vice versa.
 The changes within participant's answers from one to the other week was high, with 52/64 on average. 
 From week 1 to week 2, an increase with respect to CL (19/20, 6\;\%), basic (18/20, 6\;\%), contraposition (15/20, 5\;\%)
 and a decrease on generalization (15/20, 8\;\%) and no change on abduction (10/20, 0\;\%) answers could be observed.

@@

We identified three different clusters: 
The last two columns show the classification of the participants as deliberative, intuitive or intermediate. 
% The participants in the deliberative cluster here corresponds exactly to~\cite{khemlani:2016}. 
Our classification only differs wrt Johnson-Laird and Khemlani's 
classification~\cite{khemlani:2016} for the intermediate and intuitive cluster with respect to three participants (1, 11 and 19).
A possible explanation for this difference, could be that they
did not take into account the participants' changes between week 1 and week 2.

@@

We have investigated the cognitive adequacy of clustering by principles under WCS with respect to the individual answer patterns reported by Johnson-Laird and Khemlanid~\cite{khemlani:2016}.
Their approach is the only one we are aware of that has evaluated predictions on human syllogistic reasoning
with respect to the non-aggregated data.
However in contrast to here, they did not distinguish between the participants' responses in week 1 and week 2.

@@

% Static clusters for each participant might not be appropriate, rather, a sequential application of cognitive principles seems to be more likely
For the future, we need to investigate whether the sequential application of the cognitive principles 
can give us more insights on the participants' reasoning pattern. In particular, our hypothesis is that these principles might correspond to some of the reasoning steps within the insight model
by Johnson-Laird and Wason~\cite{johnsonlaird:1970} for the selection task~\cite{wason:68}. Similarly, a meta-analysis of the individual response patterns for the selection task has been suggested by Ragni, Kola and Johnson-Laird~\cite{ragni:2017}.
This in turn might give us a new starting point to investigate whether the cognitive principles  specified here can characterize participants' conclusions in other reasoning episodes.
